---
title: "Bottom-up mediation of microbial communities through phage infection"
author: "Larsen ML, Campbell, CE, SW Wilhelm, JT Lennon (2015)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
fontsize: 11pt
geometry: margin = 0.75in
---
#Project Summary

**Overview**: This project serves to analyze the changes in microbial chemostat populations and communities as a result of increasing trophic complexity. 
Data for this project was collected in part by MLL from a 2x2 factorial chemostat setup and by CEC who generated the microbial community sequences.

**Collaborators**

**Project Questions**

\newpage
\tableofcontents
\newpage


Data analysis for this project was completed with mothur version 1.36.1 (July 2015) and `r sessionInfo()$R.version$version.string`.





```{r setup, echo=FALSE, results='hide', message = FALSE, warning = FALSE}
# Setup Work Environment
rm(list = ls())

setwd("C:/users/meglarse/github/larsen-dissertation/chpt2_cstat-comms/community-sequencing")

# Load packages, source code, and dependencies
require(vegan);require(reshape)
require(nlme);require(ggplot2)
#require(picante); require(ape)
#require(seqinr);require(fossil)
#require(simba)

# Load data files
avg <- read.csv("./data/cstat-means.csv")
flow.counts <- read.csv("./data/20150624-popcounts.csv", header = T)
sync <- read.csv("./data/cstat-scopecounts.csv", header = T)

theme_set(theme_bw())
```


#Time-series analysis
## Summary of major results
 1. Total density did not shift through time between NL and PL -Ph treatments
 2. Total density, however, changed through time between NL and PL +Ph treatments

```{r,results = 'hide', echo = FALSE}
#create population means
flow.counts$time2 <- as.factor(flow.counts$time)
#m1 <- melt(flow.counts, id = c(trt,lim,cID,time))
m1 <- melt(flow.counts)

sem <- function(x){
  sd(x)/sqrt(length(x))
}

c1 <- cast(m1, trt + lim + time2 ~ variable, mean)
c2 <- cast(m1, trt + lim + time2 ~ variable, sem)
c1$time2 <- as.numeric(as.character(c1$time))

```

```{r AvgPopDynamics-control}

pdf(file = "./supporting-files/figures/pub-fig_chpt2-Cpopulationdynamics.pdf", 
     height = 2.5, width = 5, pointsize = 7, family="sans")

day.start <- -10
day.end <- 170

par(mfrow = c(1,2))
par(mar = c(4,6,3,0)+0.1)

# Control plots
plot(avg$day,avg$NmeanC,
       xlim = c(day.start,day.end),ylim = c(1*10^5,10^9), log = "y", xaxt = 'n',yaxt = 'n',
       xlab = "Time (d)", ylab = expression(paste("Abundance (mL"^"-1",")")), 
     cex.lab = 1.5, font.lab = 2,
       lty = 1, lwd = 1, type = "o", col = "white", bg = "white", 
       pch = 21, cex = 2)
  
  axis(1, cex.axis = 1.2,c(-120,-80,-40,0,40,80,120,160))
  ticks <- seq(5, 9, by=1)
  labels <- sapply(ticks, function(i) as.expression(bquote(10^ .(i))))
  axis(2, cex.axis = 1.2, at = c(10^5, 10^6, 10^7, 10^8, 10^9), 
       labels = labels, las = 1)
  #abline(v = 0, col = "grey", lty = 2, lwd = 2)
  
# Add error bars for the microscope counts  
#  arrows(avg$day, avg$NmeanC - avg$NseC,
#         avg$day, avg$NmeanC + avg$NseC,
#         code = 0,lwd = 2, col = "grey")

  
  if(day.start < 0){
    abline(v = 0, lwd = 1, lty = 2, col = "grey66")
  }
  points(avg$day,avg$NmeanC,
         pch = 21, col = "grey",bg = "white",
         type = 'o',lwd = 1,cex = 1, lty = 1)
  points(c1$time[c1$trt == "C" & c1$lim == "N"], 
         c1$flow.het[c1$trt == "C" & c1$lim == "N"],
         pch = 24, col = "black", bg = "green",
         type = "o", lwd = 1, cex = 1.2, lty = 1)
#  points(c1$time2[c1$trt == "C" & c1$lim == "N"], 
#         c1$flow.syn[c1$trt == "C" & c1$lim == "N"],
#         pch = 21, col = "black", bg = "black",
#         type = "o", lwd = 2, cex = 1.2, lty = 1)
box(lwd = 1.5)
text("a",cex=2.5,x=162,y=10^8.8)


par(mar = c(4,4,3,2)+0.1)

#PL
plot(avg$day,avg$PmeanC,
       xlim = c(day.start,day.end),ylim = c(1*10^5,10^9), log = "y", xaxt = 'n',yaxt = 'n',
       xlab = "Time (d)", ylab = NA,cex.lab = 1.5, 
       lty = 1, lwd = 1, type = "o", col = "grey", bg = "white", 
       pch = 21, cex = 1, font.lab = 2)
  
  axis(1, cex.axis = 1.2,c(-120,-80,-40,0,40,80,120,160))
  ticks <- seq(5, 9, by=1)
  labels <- sapply(ticks, function(i) as.expression(bquote(10^ .(i))))
  axis(2, cex.axis = 1.25, at = c(10^5, 10^6, 10^7, 10^8, 10^9), 
       labels = labels, las = 1)
  #abline(v = 0, col = "grey", lty = 2, lwd = 2)
    
# Add error bars for the microscope counts  
#  arrows(avg$day, avg$PmeanC - avg$PseC,
#         avg$day, avg$PmeanC + avg$PseC,
#         code = 0,lwd = 2, col = "grey")

  
#add a vertical line at the day of phage addition
  if(day.start < 0){
    abline(v = 0, lwd = 1, lty = 2, col = "grey66")
  }
  
  points(avg$day,avg$PmeanC,
         pch = 21, col = "grey",bg = "white",
         type = 'o',lwd = 1,cex = 1, lty = 1)
  arrows(c1$time[c1$trt == "C" & c1$lim == "P"], 
         c1$flow.het[c1$trt == "C" & c1$lim == "P"] - c2$flow.het[c2$trt == "C" & c2$lim == "P"],
         c1$time[c1$trt == "C" & c1$lim == "P"], 
         c1$flow.het[c1$trt == "C" & c1$lim == "P"] + c2$flow.het[c2$trt == "C" & c2$lim == "P"],
         code = 0,lwd = 1, col = "black")
  
    points(c1$time2[c1$trt == "C" & c1$lim == "P"], 
         c1$flow.het[c1$trt == "C" & c1$lim == "P"],
         pch = 24, col = "black", bg = "green",
         type = "o", lwd = 1, cex = 1.2, lty = 1)
#  points(c1$time2[c1$trt == "C" & c1$lim == "P"], 
#         c1$flow.syn[c1$trt == "C" & c1$lim == "P"],
#         pch = 21, col = "black", bg = "black",
#         type = "o", lwd = 2, cex = 1.2, lty = 1)

  legend("bottomright",
       c("heterotrophs","host"), 
       pch = c(24,21,22), lty = 1, 
       pt.bg = c("green","white"), 
       col = c("black","grey"),
       bty = 'n', cex = 0.8, pt.cex = 1)
box(lwd=1.5)
text("b",cex=2.5,x=162,y=10^8.8)
dev.off()
```

\newpage

```{r AvergeDynamics-treatment}  
pdf(file = "./supporting-files/figures/pub-fig_chpt2-Tpopulationdynamics.pdf", 
    height = 2.5, width = 5, pointsize = 7)

# Treatment plots
day.start <- -10
day.end <- 170

par(mfrow = c(1,2))
par(mar = c(4,6,3,0)+0.1)

plot(avg$day,avg$NmeanSI,
       xlim = c(day.start,day.end),ylim = c(1*10^5,10^9), log = "y", xaxt = 'n',yaxt = 'n',
       xlab = "Time (d)", ylab = expression(paste("Abundance (mL"^"-1",")")), 
     cex.lab = 1.5, font.lab = 2, main = NA,
       lty = 1, lwd = 1, type = "o", col = "grey", bg = "white", 
       pch = 21, cex = 1)
  
  axis(1, cex.axis = 1.2,c(-120,-80,-40,0,40,80,120,160), font = 1)
  ticks <- seq(5, 9, by=1)
  labels <- sapply(ticks, function(i) as.expression(bquote(10^ .(i))))
  axis(2, cex.axis = 1.2, at = c(10^5, 10^6, 10^7, 10^8, 10^9), 
       labels = labels, las = 1, font = 1)
  #abline(v = 0, col = "grey", lty = 2, lwd = 2)
  
# Add error bars for the microscope counts  
#  arrows(avg$day, avg$NmeanSI - avg$NseSI,
#         avg$day, avg$NmeanSI + avg$NseSI,
#         code = 0,lwd = 2, col = "grey")
#  arrows(avg$day, avg$NmeanP - avg$NseP,
#         avg$day, avg$NmeanP + avg$NseP,
#         code = 0,lwd = 2, col = "grey")

  if(day.start < 0){
    abline(v = 0, lwd = 1, lty = 2, col = "grey66")
  }
  points(avg$day,avg$NmeanSI,
         pch = 21, col = "grey",bg = "white",
         type = 'o',lwd = 1,cex = 1, lty = 1)
  points(avg$day,avg$NmeanP,
         pch = 22, col = "grey",bg = "grey80",
         type = 'o',lwd = 1,cex = 1, lty = 1)
  arrows(c1$time[c1$trt == "T" & c1$lim == "N"], 
         c1$flow.het[c1$trt == "T" & c1$lim == "N"] - c2$flow.het[c2$trt == "T" & c2$lim == "N"],
         c1$time[c1$trt == "T" & c1$lim == "N"], 
         c1$flow.het[c1$trt == "T" & c1$lim == "N"] + c2$flow.het[c2$trt == "T" & c2$lim == "N"],
         code = 0,lwd = 1, col = "black")
  
  points(c1$time[c1$trt == "T" & c1$lim == "N"], 
         c1$flow.het[c1$trt == "T" & c1$lim == "N"],
         pch = 24, col = "black", bg = "green",
         type = "o", lwd = 1, cex = 1.2, lty = 1)

  box(lwd = 1.5)
text("a",cex=2.5,x=162,y=10^8.8)


par(mar = c(4,4,3,2)+0.1)
#PL

plot(avg$day,avg$PmeanSI,
       xlim = c(day.start,day.end),ylim = c(1*10^5,10^9), log = "y", xaxt = 'n',yaxt = 'n',
       xlab = "Time (d)", ylab = NA, main = NA, cex.lab = 1.5, 
       lty = 1, lwd = 1, type = "o", col = "grey", bg = "white", 
       pch = 21, cex = 1, font.lab = 2)
  
  axis(1, cex.axis = 1.2,c(-120,-80,-40,0,40,80,120,160))
  ticks <- seq(5, 9, by=1)
  labels <- sapply(ticks, function(i) as.expression(bquote(10^ .(i))))
  axis(2, cex.axis = 1.25, at = c(10^5, 10^6, 10^7, 10^8, 10^9), 
       labels = labels, las = 1, font = 2)

# Add error bars for the microscope counts  
#  arrows(avg$day, avg$PmeanSI - avg$PseSI,
#         avg$day, avg$PmeanSI + avg$PseSI,
#         code = 0,lwd = 2, col = "grey")
#  arrows(avg$day, avg$PmeanP - avg$PseP,
#         avg$day, avg$PmeanP + avg$PseP,
#         code = 0,lwd = 2, col = "grey")
  
#add a vertical line at the day of phage addition
  if(day.start < 0){
    abline(v = 0, lwd =1, lty = 2, col = "grey66")
  }
  
  points(avg$day,avg$PmeanSI,
         pch = 21, col = "grey",bg = "white",
         type = 'o',lwd = 1,cex = 1, lty = 1)
  points(avg$day,avg$PmeanP,
         pch = 22, col = "grey",bg = "grey80",
         type = 'o',lwd = 1,cex = 1, lty = 1)
  arrows(c1$time[c1$trt == "T" & c1$lim == "P"], 
         c1$flow.het[c1$trt == "T" & c1$lim == "P"] - c2$flow.het[c2$trt == "T" & c2$lim == "P"],
         c1$time[c1$trt == "T" & c1$lim == "P"], 
         c1$flow.het[c1$trt == "T" & c1$lim == "P"] + c2$flow.het[c2$trt == "T" & c2$lim == "P"],
         code = 0,lwd = 1, col = "black")
  
    points(c1$time[c1$trt == "T" & c1$lim == "P"], 
         c1$flow.het[c1$trt == "T" & c1$lim == "P"],
         pch = 24, col = "black", bg = "green",
         type = "o", lwd = 1, cex = 1.2, lty = 1)
 
box(lwd=1.5)
text("b",cex=2.5,x=162,y=10^8.8)
#par(new=TRUE)
#plot(avg$day,vbP,
#     type = "l", lty = 1, lwd = 2, col = "red",
#     xaxt="n",yaxt="n",xlab="",ylab="")
#axis(4, las = 1, col = "red", cex.axis = 1.25)
#mtext("Phage to Cyanobacteria",side=4,line=3)

legend("bottomright",
       c("heterotrophs","host","phage"), 
       pch = c(24,21,22), lty = 1, 
       pt.bg = c("green","white","grey"), 
       col = c("black","grey","grey"),
       bty = 'n', cex = 0.8, pt.cex = 1)

dev.off()
```

\newpage
## RMANOVA
```{r RMANOVA}
#Q: Does the het density change through time? 
#Q: Is it correlated with the Synechococcus density?

control <- flow.counts[flow.counts$trt == "C",]
treatment <- flow.counts[flow.counts$trt == "T",]
NL <- flow.counts[flow.counts$lim == "N",]
PL <- flow.counts[flow.counts$lim == "P",]

ar1.hetC <- lme(flow.het ~ lim * time , random = ~ 1 | cID, 
            correlation = corAR1(form = ~ 1 | cID),
            data = control)

ar1.hetT <- lme(flow.het ~ lim * time , random = ~ 1 | cID, 
            correlation = corAR1(form = ~ 1 | cID),
            data = treatment)
ar1.hetN <- lme(flow.het ~ trt * time , random = ~ 1 | cID, 
            correlation = corAR1(form = ~ 1 | cID),
            data = NL)
ar1.hetP <- lme(flow.het ~ trt * time , random = ~ 1 | cID, 
            correlation = corAR1(form = ~ 1 | cID),
            data = PL)

#arma.het <- lme(flow.het ~ lim * time * trt, random = ~ 1 | cID, 
#            correlation = corARMA(form = ~ 1 , p = 1,q = 2),
#            data = flow.counts)

#anova(ar1.het,arma.het)
summary(ar1.het)
anova.lme(ar1.hetC, type = "sequential", adjustSigma = F)
anova.lme(ar1.hetT, type = "sequential", adjustSigma = F)
anova.lme(ar1.hetN, type = "sequential", adjustSigma = F)
anova.lme(ar1.hetP, type = "sequential", adjustSigma = F)
```

```{r CCF}

CN <- flow.counts[flow.counts$trt=="C" & flow.counts$lim=="N"]
TN <- flow.counts[flow.counts$trt=="T" & flow.counts$lim=="N"]
CP <- flow.counts[flow.counts$trt=="C" & flow.counts$lim=="P"]
TP <- flow.counts[flow.counts$trt=="T" & flow.counts$lim=="P"]

pdf(file = "./supporting-files/figures/pub-fig_C-CCF.pdf", width = 6, height = 3, pointsize = 7)
par(mfrow =c(1,2))
ccf.CN <- ccf(CN$flow.syn, CN$flow.het, lag.max = 6,main = "NL, -Ph")
ccf.CP <- ccf(CP$flow.syn, CP$flow.het, lag.max = 6, main = "PL, -Ph")
dev.off()

pdf(file = "./supporting-files/figures/pub-fig_T-CCF.pdf", width = 6, height = 3, pointsize = 7)
par(mfrow =c(1,2))
ccf.TN <- ccf(TN$flow.syn, TN$flow.het, lag.max = 6, main = "NL, +Ph")
ccf.TP <- ccf(TP$flow.syn, TP$flow.het, lag.max = 6, main = "PL, +Ph")
dev.off()

ccf.results <- data.frame(ccf.CN$lag,
                          ccf.CN$acf,
                          ccf.TN$acf,
                          ccf.CP$acf,
                          ccf.TP$acf)
colnames(ccf.results) <- c("lag","NL Control","NL +Phage", "PL Control","PL +Phage")

#require("pander")
#pandoc.table(ccf.results, style = "rmarkdown", round = 3, 
#             caption = "acf values for each group")
```


#Community analysis
##Summary of major findings

\newpage

```{r}
# load packages and source code
require("vegan")
source("./bin/DiversityFunctions.r")  
#source("./bin/MothurTools.R")

# load required functions
se <- function(x){sd(x)/sqrt(length(x))}                                   

# data file information
level <- "unique"
shared <- "./mothur/20151029/Cstat_all.final.an.shared"
design.in <- "./data/design.txt"
plot.title <- "Cstat_prelim"
```

```{r 2.2_preprocessing, echo = FALSE}
# Import Site by OTU Matrix
cstat_data <- t(read.otu(shared, level))             
design <- read.delim(design.in, header=T, row.names=1)

# Remove Zero Sum OTUs -- singletons and doubletons
cstat_data <- cstat_data[!(rowSums(abs(cstat_data)) < 2),] 
#dim(cstat_data)

# Remove cyanobacteria calls
rm.rows <- c(17,23,28,31,36,38,40,46,52)
cstat_data <- cstat_data[-rm.rows,]
cstat_data_syn <- cstat_data
#dim(cstat_data)
```

```{r 2.2_beta_diversity, echo = FALSE , fig.height=5, fig.width=5}
# remove synechococcus for beta diversity analysis
syn <- cstat_data[3,]
cstat_data <- cstat_data[-3,]

# Calculate Presense Absence
dataPA <- (cstat_data > 0)*1 

# Calculating Relative Abundance
dataREL <- cstat_data
  for(i in 1:ncol(cstat_data)){
    dataREL[,i] = cstat_data[,i]/sum(cstat_data[,i])
    }  

# Create Distance Matrix
sampleREL.dist <- vegdist(decostand(t(dataREL),method="log"),method="bray")

# Principal Coordinates Analysis
cstat_pcoa <- cmdscale(sampleREL.dist,k=3,eig=TRUE,add=FALSE) 
  # Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
    	# eig=TRUE returns eigenvalues; k = # of dimensions to calculate

# Percent Variance Explained Using PCoA (Axis 1,2,3)
explainvar1 <- round(cstat_pcoa$eig[1]/sum(cstat_pcoa$eig)*100,1) 
explainvar2 <- round(cstat_pcoa$eig[2]/sum(cstat_pcoa$eig)*100,1)
explainvar3 <- round(cstat_pcoa$eig[3]/sum(cstat_pcoa$eig)*100,1)
  
pcoap <- merge(as.data.frame(cstat_pcoa$points),design,by=0,all.x=T)
rownames(pcoap) <- rownames(cstat_pcoa$points)

#pcoap <- subset(pcoap,pcoap$limitation == "T")

pdf(file = "./supporting-files/figures/pub-fig_betaPCOA.pdf", width = 5, height = 5, pointsize = 10)

# Plot Parameters
par(mar=c(5,5,1,1))#, oma=c(3,1,1,1)+0.1 )
x.dim <- c(min(pcoap$V1)-(max(pcoap$V1)*0.1),max(pcoap$V1)+(max(pcoap$V1)*0.1))
y.dim <- c(min(pcoap$V2)-(max(pcoap$V2)*0.1),max(pcoap$V2)+(max(pcoap$V2)*0.1))

# Initiate Plot
plot(pcoap$V1, pcoap$V2, xlab=paste("PCoA Axis 1 (",explainvar1, "%)", sep="")
    , ylab=paste("PCoA Axis 2 (",explainvar2, "%)", sep=""), 
    xlim=x.dim, ylim=y.dim, pch=16, cex=2.0, type="n",xaxt="n",
    yaxt="n", cex.lab=1.5, cex.axis=1.2)
axis(side=1, las=1)   
axis(side=2, las=1)    
abline(h=0, lty="dotted",col = "grey50")  
abline(v=0, lty="dotted",col = "grey50")
box(lwd=2)

# Make Plot Symbols in PCoA Reflect Treatment
 
mol.shape <- rep(NA, dim(pcoap)[1])
  for (i in 1:length(mol.shape)){
    if (pcoap$treatment[i] == "T"){mol.shape[i] = 21}
    else {mol.shape[i] = 22}
	}

slope.color <- rep(NA, dim(pcoap)[1])
  for (i in 1:length(slope.color)){
    if (pcoap$limitation[i] == "N") {slope.color[i] = "black"}
    else {slope.color[i] = "white"}  
	} 

# Add Points & Ellipses
points(pcoap$V1, pcoap$V2, pch = mol.shape, cex = 3.5, 
       bg = slope.color, col = "gray25")

legend("bottomleft",pch = c(21,22, 21, 22), col = c(""))
#text(pcoap$V1, pcoap$V2, labels = pcoap$time, col = "grey50")  
#text(pcoap$V1, pcoap$V2, labels = pcoap$cID, pos = 1) 

#dev.off()

p = ggplot(pcoap, aes(x = V1, y=V2, color = treatment:limitation)) + 
  geom_hline(yintercept = 0, col = "grey85",lty = 2) +
  geom_vline(xintercept = 0, col = "grey55",lty=2) +
  geom_point(size = 5) + scale_color_brewer(palette = 4) + geom_point(size = 5, col = "grey25",pch=21) + xlab(paste("PCoA Axis 1 (",explainvar1,"%",")")) + ylab(paste("PCoA Axis 2 (",explainvar2,"%",")"))

ggsave(filename = "./supporting-files/figures/pub-fig_beta.pdf", plot = p, height = 6, width = 7, units = "in", scale = 1)  

#p = ggplot(pcoap,aes(x=V1, y = V2, shape = treatment )) + geom_point(size =4) + theme(legend.position="none")
#p = p + facet_grid(~cID) +  
#  geom_path(aes(x = V1, y = V2, size = time/1.5), color = "grey80") +
#  geom_point(aes(size = time), color = "black") + xlab("Axis.1") + ylab("Axis.2") 

#print(p)

#ggsave(filename = "./figure_cIDbeta.pdf", plot = p, height = 2.5, width = 10, units = "in", scale = 2)

```

```{r 2.2_PERMANOVA}
#BC = t(dataREL)
#BC = vegdist(t(dataREL),method="bray")
#df =  design
#results3 <- adonis(BC ~ time*treatment*limitation, data = df, 
#                  strata = df$cID, permutations = 999, "bray")

cstat.adonis.data <- t(dataREL)
cstat.adonis.data <- merge(design,cstat.adonis.data, by=0,all.x=T)
cstat.adonis.data <- cstat.adonis.data[-17,]
#cstat.adonis.data <- subset(cstat.adonis.data,cstat.adonis.data$trt == "T")
#cstat.adonis.data <- cstat.adonis.data[,1:36]

results3 <- adonis(cstat.adonis.data[,6:ncol(cstat.adonis.data)] ~ 
         limitation*treatment*time, strata = cstat.adonis.data$cID, data = cstat.adonis.data,
       method="bray", permutations=999)


```
\newpage

#PERMANOVA analyses for each treatment group
```{r 2.2_PERMANOVA-lim+Ph}
#BC = t(dataREL)
#BC = vegdist(t(dataREL),method="bray")
#df =  design
#results3 <- adonis(BC ~ time*treatment*limitation, data = df, 
#                  strata = df$cID, permutations = 999, "bray")

cstat.adonis.data <- t(dataREL)
designT <- design[design$treatment == "T",]
cstat.adonis.data <- merge(designT,cstat.adonis.data, by=0,all.x=T)
cstat.adonis.data <- cstat.adonis.data[-12,]


results3 <- adonis(cstat.adonis.data[,6:ncol(cstat.adonis.data)] ~ 
         limitation*day, strata = cstat.adonis.data$cID, data = cstat.adonis.data,
       method="bray", permutations=999)


```

```{r 2.2_PERMANOVA-lim-Ph}
#BC = t(dataREL)
#BC = vegdist(t(dataREL),method="bray")
#df =  design
#results3 <- adonis(BC ~ time*treatment*limitation, data = df, 
#                  strata = df$cID, permutations = 999, "bray")

cstat.adonis.data <- t(dataREL)
designC <- design[design$treatment == "C",]
cstat.adonis.data <- merge(designC,cstat.adonis.data, by=0,all.x=T)


results3 <- adonis(cstat.adonis.data[,6:ncol(cstat.adonis.data)] ~ 
         limitation*day, strata = cstat.adonis.data$cID, data = cstat.adonis.data,
       method="bray", permutations=999)


```

```{r 2.2_PERMANOVA-NL}
#BC = t(dataREL)
#BC = vegdist(t(dataREL),method="bray")
#df =  design
#results3 <- adonis(BC ~ time*treatment*limitation, data = df, 
#                  strata = df$cID, permutations = 999, "bray")

cstat.adonis.data <- t(dataREL)
designNL <- design[design$limitation == "N",]
cstat.adonis.data <- merge(designNL,cstat.adonis.data, by=0,all.x=T)


results3 <- adonis(cstat.adonis.data[,6:ncol(cstat.adonis.data)] ~ 
         treatment*day, strata = cstat.adonis.data$cID, data = cstat.adonis.data,
       method="bray", permutations=999)


```

```{r 2.2_PERMANOVA-PL}
#BC = t(dataREL)
#BC = vegdist(t(dataREL),method="bray")
#df =  design
#results3 <- adonis(BC ~ time*treatment*limitation, data = df, 
#                  strata = df$cID, permutations = 999, "bray")

cstat.adonis.data <- t(dataREL)
designPL <- design[design$limitation == "P",]
cstat.adonis.data <- merge(designPL,cstat.adonis.data, by=0,all.x=T)
cstat.adonis.data <- cstat.adonis.data[-9,]

results3 <- adonis(cstat.adonis.data[,6:ncol(cstat.adonis.data)] ~ 
         treatment*day, strata = cstat.adonis.data$cID, data = cstat.adonis.data,
       method="bray", permutations=999)
```

```{r 2.2_temporal_composition, echo = FALSE, fig.width=7.5, fig.height=4}
library(ggplot2)
library(ggthemes)
library(reshape2)

# Select most abundant taxa with synechococcus (OTU3); 
cpNo <- c(3.57,2.33,1.85,3.08,5.33,4.86,2.50,1)
cstat_data_top <- data.frame(t(cstat_data[1:7,]),syn)
#cstat_data_top$others <- NA

#for(i in 1:nrow(cstat_data_top)){
#  cstat_data_top[i,8] <- rowSums(t(abs(cstat_data[8:nrow(cstat_data),])))[[i]]
#}  

#adjust number of reads by mean copy number for each strain
#for(i in 1:length(cpNo)){
#  cstat_data_top[,i] <- cstat_data_top[,i]/cpNo[i]
#}

cstat_data_top <- t(cstat_data_top)

# With Syn
dataRELtop <- cstat_data_top
  for(i in 1:ncol(cstat_data_top)){
    dataRELtop[,i] = cstat_data_top[,i]/sum(cstat_data_top[,i])
    } 

top.abd.design <- merge(design,t(dataRELtop), by=0,all.x=T)

colnames(top.abd.design) <- c(colnames(top.abd.design)[1:6],"Sulfitobacter","Sphingomonas",
                              "Rhizobium","Arthrobacter","Alcanivorax","Pseudomonas","others","Synechococcus")

write.csv(top.abd.design, file = "./supporting-files/data/arima-proc-rel.csv",row.names = FALSE)


melt.abd <- melt(top.abd.design, id=c("time","day","cID","Row.names","limitation","treatment"))
melt.abd$day <- as.numeric(as.character(melt.abd$day))



tes <- melt.abd[-which(melt.abd$cID == "PL3"),]
colnames(tes)<- c(colnames(tes)[1:6],"Taxa","value")
p <- ggplot(tes,aes(x = day, y = value, fill = Taxa)) + 
  geom_area(position = 'stack', colour ="black") + 
  facet_wrap(~cID, ncol = 4) +
  #facet_gridtreatment ~ limitation) +
  labs(x = "Time (days)", y = "Relative Abundance")+
  #scale_fill_grey(start = 0 , end = 0.9)
  scale_fill_brewer()
  
#p$data$Taxa <- factor(p$data$Taxa, levels = #c("Pseudomonas","Alcanivorax","Arthrobacter","Rhizobium","Sphingomonas","Sulfitobacter"))

print(p)

ggsave(filename="./supporting-files/figures/pub-fig_taxarelabd-cid-syn.pdf", plot = p,width = 8, height = 4, units = "in",scale = 1.5)
```

```{r}
# Without Syn
cstat_data_top_wosyn <- cstat_data_top[-8,]
dataRELtop_wo <- cstat_data_top_wosyn
  for(i in 1:ncol(cstat_data_top_wosyn)){
    dataRELtop_wo[,i] = cstat_data_top_wosyn[,i]/sum(cstat_data_top_wosyn[,i])
    } 
relAbd <- merge(design,t(dataRELtop_wo), by=0,all.x=T)
relAbdflow <- merge(relAbd,flow.counts, by.x = "Row.names", by.y = "seq.ID")
relAbdflow <- relAbdflow[,c(1:13,18:19)]

for(i in 1:nrow(relAbdflow)){
  relAbdflow[i,7:13] <- relAbdflow[i,7:13]*relAbdflow[i,15]
}

relabdflow2 <- relAbdflow[,1:12]
relabdflow2 = merge(relabdflow2,sync,by="day")


colnames(relAbd) <- c(colnames(relAbd)[1:6],"Sulfitobacter","Sphingomonas",
                              "Rhizobium","Arthrobacter","Alcanivorax","Pseudomonas","008")
melt.abd <- melt(relAbd, id=c("time","day","cID","Row.names","limitation","treatment"))
melt.abd$day <- as.numeric(as.character(melt.abd$day))



tes <- melt.abd[-which(melt.abd$cID == "PL3"),]
colnames(tes)<- c(colnames(tes)[1:6],"Taxa","value")

# as cID
p2 <- ggplot(tes,aes(x = day, y = value, fill = Taxa)) + 
  geom_area(position = 'stack', colour ="black") + 
  facet_wrap(~cID, ncol = 4) +
  #facet_gridtreatment ~ limitation) +
  labs(x = "Time (days)", y = "Relative Abundance")+
  #scale_fill_grey(start = 0 , end = 0.9)
  scale_fill_brewer()
print(p2)


```

```{r}
colnames(relAbdflow) <- c(colnames(relAbdflow)[1:6],"Sulfitobacter","Sphingomonas",
                              "Rhizobium","Arthrobacter","Alcanivorax","Pseudomonas","008","flow.syn","flow.het")
melt.abd <- melt(relAbdflow, id=c("time.x","day","cID.x","Row.names","limitation","treatment"))
melt.abd$day <- as.numeric(as.character(melt.abd$day))



tes <- melt.abd[-which(melt.abd$cID == "PL3"),]
colnames(tes)<- c(colnames(tes)[1:6],"Taxa","value")
#taxa <- unique(tes$Taxa)[1:6]
#tes = tes[tes$Taxa == taxa,]

p3 <- ggplot(tes,aes(x = day, y = value, fill = Taxa)) + 
  stat_summary(fun.y = "mean",geom="line")+
  geom_area(position = 'stack', colour ="black") + 
  facet_wrap(trt ~ lim) +
  #facet_gridtreatment ~ limitation) +
  labs(x = "Time (days)", y = "Relative Abundance")+
  #scale_fill_grey(start = 0 , end = 0.9)
  scale_fill_brewer()
print(p3)
```

```{r perm-tax}

dat = p2$data
#dat$time = as.factor(dat$time)
for(i in unique(dat$Taxa)){
  tmpdat = dat[dat$Taxa == i,]
  
  fit <- lme(value ~ limitation * time * treatment, random = ~ 1 | cID,
                 correlation = corAR1(form = ~ 1 | cID),
                 data = tmpdat, na.action = na.exclude)
  fit.dat = anova.lme(fit, type = "marginal", adjustSigma = F)
  print(i)
  print(fit.dat)
  
}

# calculate the mean relative densities for each of the major taxa for each treatment
dat.rhi <- dat[dat$Taxa == "Sulfitobacter",]
aggregate(value~ limitation + treatment, data = dat.rhi, mean)
aggregate(value~ limitation + treatment, data = dat.rhi, sd)
aggregate(value~ cID, data = dat.rhi, mean, na.action = na.exclude)
aggregate(value~ limitation, data = dat.rhi, mean)
aggregate(value~ limitation, data = dat.rhi, sd)

```


```{r}

#install.packages("Hmisc")
require(Hmisc)

mat.N <- as.matrix(top.abd.design[,7:13][top.abd.design$limitation=="N",])
mat.P <- as.matrix(top.abd.design[,7:13][top.abd.design$limitation=="P",])
mat.P <- na.omit(mat.P)
ts.cor.N <- rcorr(mat.N,type="pearson")
ts.cor.P <- rcorr(mat.P,type="pearson")

syn.N <- mat.N[,3]
syn.P <- mat.P[,3]
sulf.N <- mat.N[,1]
sph.N <- mat.N[,2]
ath.N <- mat.N[,5]
ath.N <- mat.N[,5]

for(i in unique(top.abd.design$cID)[-9]){
  tmp <- as.matrix(top.abd.design[top.abd.design$cID == i,7:13])
  tmp.cor <- rcorr(tmp,type="pearson")
  print(i)
  print(tmp.cor)
}
```
```{r}
pdf(file = "./figure_NLccf.pdf", width = 5, height = 4.5, pointsize= 10)

par(mfrow = c(2,4))
for(i in 1:7){
  ccf(syn.N,mat.N[,i], main = colnames(mat.N)[i])
}

dev.off()
#Sphingomonas = out of phase, Rhizobium out of phase, Athrobacter
```

```{r}
pdf(file = "./figure_PLccf.pdf", width = 5, height = 4.5, pointsize = 10)


par(mfrow = c(2,4))
for(i in 1:7){
  ccf(syn.P,mat.P[,i], main = colnames(mat.P)[i])
}

dev.off()
#Sulfitobacter out of phase. lag -2, Rhizobium, 
```
