---
title: "Bottom-up mediation of microbial communities through phage infection"
author: "Larsen ML, Campbell, CE, SW Wilhelm, JT Lennon (2015)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
fontsize: 11pt
geometry: margin = 0.75in
---

**Overview**: This project serves to analyze the changes in microbial chemostat populations and communities as a result of increasing trophic complexity. 
Data for this project was collected in part by MLL from a 2x2 factorial chemostat setup and by CEC who generated the microbial community sequences.

Data analysis for this project was completed with mothur version 1.36.1 (July 2015) and `r sessionInfo()$R.version$version.string`.

```{r, echo=FALSE, results='hide', message = FALSE, warning = FALSE}
# Setup Work Environment
rm(list = ls())
#setwd("C:/Users/Megan/GitHub/Cstat_Communities")
#setwd("C:/Users/Megan/GitHub/larsen-dissertation/chpt2_cstat-comms/community-sequencing")
setwd("C:/users/meglarse/github/larsen-dissertation/chpt2_cstat-comms/community-sequencing")

# Load packages, source code, and dependencies
require(vegan);require(reshape)
require(nlme);require(ggplot2)
#require(picante); require(ape)
#require(seqinr);require(fossil)
#require(simba)

# Load data files
avg <- read.csv("./data/cstat-means.csv")
flow.counts <- read.csv("./data/20150624-popcounts.csv", header = T)
sync <- read.csv("./data/cstat-scopecounts.csv", header = T)

theme_set(theme_bw())
```


##1 FIGURE 1: Temporal dynamics
```{r,results = 'hide', echo = FALSE}
#create population means
flow.counts$time2 <- as.factor(flow.counts$time)
#m1 <- melt(flow.counts, id = c(trt,lim,cID,time))
m1 <- melt(flow.counts)

sem <- function(x){
  sd(x)/sqrt(length(x))
}

c1 <- cast(m1, trt + lim + time2 ~ variable, mean)
c2 <- cast(m1, trt + lim + time2 ~ variable, sem)
c1$time2 <- as.numeric(as.character(c1$time))
avg <- read.csv("./data/cstat-means.csv")
```

```{r, fig.height=6, fig.width=8, echo=FALSE, fig.cap="Population Dynamics"}
# Suggested changes

#png(filename = "./figures/time-series.png",
#    width = 1200, height = 1000, units = "px", pointsize = 20,
#   bg = "white", family = "sans", type = "cairo")

#pdf(file = "./figure_populationdynamics.pdf",height = 4.5, width = 5, pointsize = 8)

day.start <- -10
day.end <- 170

par(mfrow = c(2,2))
par(mar = c(2,6,3,1)+0.1)

# Control plots
plot(avg$day,avg$NmeanC,
       xlim = c(day.start,day.end),ylim = c(1*10^5,10^9), log = "y", xaxt = 'n',yaxt = 'n',
       xlab = NA, ylab = expression(paste("Abundance (mL"^"-1",")")), 
     cex.lab = 1.5, font.lab = 2, main = "N-Limited",
       lty = 1, lwd = 1, type = "o", col = "white", bg = "white", 
       pch = 21, cex = 2)
  
  #axis(1, cex.axis = 1.2,c(-120,-80,-40,0,40,80,120,160))
  ticks <- seq(5, 9, by=1)
  labels <- sapply(ticks, function(i) as.expression(bquote(10^ .(i))))
  axis(2, cex.axis = 1.2, at = c(10^5, 10^6, 10^7, 10^8, 10^9), 
       labels = labels, las = 1)
  #abline(v = 0, col = "grey", lty = 2, lwd = 2)
  
# Add error bars for the microscope counts  
#  arrows(avg$day, avg$NmeanC - avg$NseC,
#         avg$day, avg$NmeanC + avg$NseC,
#         code = 0,lwd = 2, col = "grey")

  
  if(day.start < 0){
    abline(v = 0, lwd = 1, lty = 2, col = "grey66")
  }
  points(avg$day,avg$NmeanC,
         pch = 21, col = "grey",bg = "white",
         type = 'o',lwd = 1,cex = 1, lty = 1)
  points(c1$time[c1$trt == "C" & c1$lim == "N"], 
         c1$flow.het[c1$trt == "C" & c1$lim == "N"],
         pch = 24, col = "black", bg = "green",
         type = "o", lwd = 1, cex = 1.2, lty = 1)
#  points(c1$time2[c1$trt == "C" & c1$lim == "N"], 
#         c1$flow.syn[c1$trt == "C" & c1$lim == "N"],
#         pch = 21, col = "black", bg = "black",
#         type = "o", lwd = 2, cex = 1.2, lty = 1)
box(lwd = 2)

par(mar = c(2,4,3,3)+0.1)
#PL
plot(avg$day,avg$PmeanC,
       xlim = c(day.start,day.end),ylim = c(1*10^5,10^9), log = "y", xaxt = 'n',yaxt = 'n',
       xlab = NA, ylab = NA, main = "P-Limited",cex.lab = 1.5, 
       lty = 1, lwd = 1, type = "o", col = "grey", bg = "white", 
       pch = 21, cex = 1, font.lab = 2)
  
  #axis(1, cex.axis = 1.2,c(-120,-80,-40,0,40,80,120,160))
  ticks <- seq(5, 9, by=1)
  labels <- sapply(ticks, function(i) as.expression(bquote(10^ .(i))))
  axis(2, cex.axis = 1.25, at = c(10^5, 10^6, 10^7, 10^8, 10^9), 
       labels = labels, las = 1)
  #abline(v = 0, col = "grey", lty = 2, lwd = 2)
    
# Add error bars for the microscope counts  
#  arrows(avg$day, avg$PmeanC - avg$PseC,
#         avg$day, avg$PmeanC + avg$PseC,
#         code = 0,lwd = 2, col = "grey")

  
#add a vertical line at the day of phage addition
  if(day.start < 0){
    abline(v = 0, lwd = 1, lty = 2, col = "grey66")
  }
  
  points(avg$day,avg$PmeanC,
         pch = 21, col = "grey",bg = "white",
         type = 'o',lwd = 1,cex = 1, lty = 1)
  arrows(c1$time[c1$trt == "C" & c1$lim == "P"], 
         c1$flow.het[c1$trt == "C" & c1$lim == "P"] - c2$flow.het[c2$trt == "C" & c2$lim == "P"],
         c1$time[c1$trt == "C" & c1$lim == "P"], 
         c1$flow.het[c1$trt == "C" & c1$lim == "P"] + c2$flow.het[c2$trt == "C" & c2$lim == "P"],
         code = 0,lwd = 1, col = "black")
  
    points(c1$time2[c1$trt == "C" & c1$lim == "P"], 
         c1$flow.het[c1$trt == "C" & c1$lim == "P"],
         pch = 24, col = "black", bg = "green",
         type = "o", lwd = 1, cex = 1.2, lty = 1)
#  points(c1$time2[c1$trt == "C" & c1$lim == "P"], 
#         c1$flow.syn[c1$trt == "C" & c1$lim == "P"],
#         pch = 21, col = "black", bg = "black",
#         type = "o", lwd = 2, cex = 1.2, lty = 1)

  
  box(lwd=2)


# Treatment plots
par(mar = c(5,6,1,1)+0.1)

plot(avg$day,avg$NmeanSI,
       xlim = c(day.start,day.end),ylim = c(1*10^5,10^9), log = "y", xaxt = 'n',yaxt = 'n',
       xlab = "Time (d)", ylab = expression(paste("Abundance (mL"^"-1",")")), 
     cex.lab = 1.5, font.lab = 2, main = NA,
       lty = 1, lwd = 1, type = "o", col = "grey", bg = "white", 
       pch = 21, cex = 1)
  
  axis(1, cex.axis = 1.2,c(-120,-80,-40,0,40,80,120,160), font = 1)
  ticks <- seq(5, 9, by=1)
  labels <- sapply(ticks, function(i) as.expression(bquote(10^ .(i))))
  axis(2, cex.axis = 1.2, at = c(10^5, 10^6, 10^7, 10^8, 10^9), 
       labels = labels, las = 1, font = 1)
  #abline(v = 0, col = "grey", lty = 2, lwd = 2)
  
# Add error bars for the microscope counts  
#  arrows(avg$day, avg$NmeanSI - avg$NseSI,
#         avg$day, avg$NmeanSI + avg$NseSI,
#         code = 0,lwd = 2, col = "grey")
#  arrows(avg$day, avg$NmeanP - avg$NseP,
#         avg$day, avg$NmeanP + avg$NseP,
#         code = 0,lwd = 2, col = "grey")

  if(day.start < 0){
    abline(v = 0, lwd = 1, lty = 2, col = "grey66")
  }
  points(avg$day,avg$NmeanSI,
         pch = 21, col = "grey",bg = "white",
         type = 'o',lwd = 1,cex = 1, lty = 1)
  points(avg$day,avg$NmeanP,
         pch = 22, col = "grey",bg = "grey80",
         type = 'o',lwd = 1,cex = 1, lty = 1)
  arrows(c1$time[c1$trt == "T" & c1$lim == "N"], 
         c1$flow.het[c1$trt == "T" & c1$lim == "N"] - c2$flow.het[c2$trt == "T" & c2$lim == "N"],
         c1$time[c1$trt == "T" & c1$lim == "N"], 
         c1$flow.het[c1$trt == "T" & c1$lim == "N"] + c2$flow.het[c2$trt == "T" & c2$lim == "N"],
         code = 0,lwd = 1, col = "black")
  
  points(c1$time[c1$trt == "T" & c1$lim == "N"], 
         c1$flow.het[c1$trt == "T" & c1$lim == "N"],
         pch = 24, col = "black", bg = "green",
         type = "o", lwd = 1, cex = 1.2, lty = 1)

  box(lwd = 2)

#PL
par(mar = c(5,4,1,3)+0.1)

plot(avg$day,avg$PmeanSI,
       xlim = c(day.start,day.end),ylim = c(1*10^5,10^9), log = "y", xaxt = 'n',yaxt = 'n',
       xlab = "Time (d)", ylab = NA, main = NA, cex.lab = 1.5, 
       lty = 1, lwd = 1, type = "o", col = "grey", bg = "white", 
       pch = 21, cex = 1, font.lab = 2)
  
  axis(1, cex.axis = 1.2,c(-120,-80,-40,0,40,80,120,160))
  ticks <- seq(5, 9, by=1)
  labels <- sapply(ticks, function(i) as.expression(bquote(10^ .(i))))
  axis(2, cex.axis = 1.25, at = c(10^5, 10^6, 10^7, 10^8, 10^9), 
       labels = labels, las = 1, font = 2)

# Add error bars for the microscope counts  
#  arrows(avg$day, avg$PmeanSI - avg$PseSI,
#         avg$day, avg$PmeanSI + avg$PseSI,
#         code = 0,lwd = 2, col = "grey")
#  arrows(avg$day, avg$PmeanP - avg$PseP,
#         avg$day, avg$PmeanP + avg$PseP,
#         code = 0,lwd = 2, col = "grey")
  
#add a vertical line at the day of phage addition
  if(day.start < 0){
    abline(v = 0, lwd =1, lty = 2, col = "grey66")
  }
  
  points(avg$day,avg$PmeanSI,
         pch = 21, col = "grey",bg = "white",
         type = 'o',lwd = 1,cex = 1, lty = 1)
  points(avg$day,avg$PmeanP,
         pch = 22, col = "grey",bg = "grey80",
         type = 'o',lwd = 1,cex = 1, lty = 1)
  arrows(c1$time[c1$trt == "T" & c1$lim == "P"], 
         c1$flow.het[c1$trt == "T" & c1$lim == "P"] - c2$flow.het[c2$trt == "T" & c2$lim == "P"],
         c1$time[c1$trt == "T" & c1$lim == "P"], 
         c1$flow.het[c1$trt == "T" & c1$lim == "P"] + c2$flow.het[c2$trt == "T" & c2$lim == "P"],
         code = 0,lwd = 1, col = "black")
  
    points(c1$time[c1$trt == "T" & c1$lim == "P"], 
         c1$flow.het[c1$trt == "T" & c1$lim == "P"],
         pch = 24, col = "black", bg = "green",
         type = "o", lwd = 1, cex = 1.2, lty = 1)
 
box(lwd=2)

#par(new=TRUE)
#plot(avg$day,vbP,
#     type = "l", lty = 1, lwd = 2, col = "red",
#     xaxt="n",yaxt="n",xlab="",ylab="")
#axis(4, las = 1, col = "red", cex.axis = 1.25)
#mtext("Phage to Cyanobacteria",side=4,line=3)

legend("bottomright",
       c("heterotrophs","host","phage"), 
       pch = c(24,21,22), lty = 1, 
       pt.bg = c("green","white","grey"), 
       col = c("black","grey","grey"),
       bty = 'n', cex = 0.8, pt.cex = 1)

#dev.off()
```


```{r RMANOVA}
#Q: Does the het density change through time? 
#Q: Is it correlated with the Synechococcus density?

ar1.het <- lme(flow.het ~ lim * time * trt, random = ~ 1 | cID, 
            correlation = corAR1(form = ~ 1 | cID),
            data = flow.counts)
#arma.het <- lme(flow.het ~ lim * time * trt, random = ~ 1 | cID, 
#            correlation = corARMA(form = ~ 1 , p = 1,q = 2),
#            data = flow.counts)

#anova(ar1.het,arma.het)
summary(ar1.het)
anova.lme(ar1.het, type = "marginal", adjustSigma = F)
```

```{r CCF}

CN <- flow.counts[flow.counts$trt=="C" & flow.counts$lim=="N",]
TN <- flow.counts[flow.counts$trt=="T" & flow.counts$lim=="N",]
CP <- flow.counts[flow.counts$trt=="C" & flow.counts$lim=="P",]
TP <- flow.counts[flow.counts$trt=="T" & flow.counts$lim=="P",]

par(mfrow =c(2,2))
ccf.CN <- ccf(CN$flow.syn, CN$flow.het, lag.max = 6)
ccf.TN <- ccf(TN$flow.syn, TN$flow.het, lag.max = 6)
ccf.CP <- ccf(CP$flow.syn, CP$flow.het, lag.max = 6)
ccf.TP <- ccf(TP$flow.syn, TP$flow.het, lag.max = 6)

ccf.results <- data.frame(ccf.CN$lag,
                          ccf.CN$acf,
                          ccf.TN$acf,
                          ccf.CP$acf,
                          ccf.TP$acf)
colnames(ccf.results) <- c("lag","NL Control","NL +Phage", "PL Control","PL +Phage")

require("pander")
pandoc.table(ccf.results, style = "rmarkdown", round = 3, 
             caption = "acf values for each group")
```


##2 Community Analysis
```{r 2.2_section_data}
# load packages and source code
require("vegan")
source("./bin/DiversityFunctions.r")  
#source("./bin/MothurTools.R")

# load required functions
se <- function(x){sd(x)/sqrt(length(x))}                                   

# data file information
level <- "unique"
shared <- "./mothur/20151029/Cstat_all.final.an.shared"
design.in <- "./data/design.txt"
plot.title <- "Cstat_prelim"
```

```{r 2.2_preprocessing, echo = FALSE}
# Import Site by OTU Matrix
cstat_data <- t(read.otu(shared, level))             
design <- read.delim(design.in, header=T, row.names=1)

# Remove Zero Sum OTUs -- singletons and doubletons
cstat_data <- cstat_data[!(rowSums(abs(cstat_data)) < 2),] 

# Remove cyanobacteria calls
rm.rows <- c(17,23,28,31,36,38,40,46,52)
cstat_data <- cstat_data[-rm.rows,]
cstat_data_syn <- cstat_data
#dim(cstat_data)
```

```{r 2.2_beta_diversity, echo = FALSE , fig.height=5, fig.width=5}
# remove synechococcus for beta diversity analysis
cstat_data <- cstat_data[-3,]

# Calculate Presense Absence
dataPA <- (cstat_data > 0)*1 

# Calculating Relative Abundance
dataREL <- cstat_data
  for(i in 1:ncol(cstat_data)){
    dataREL[,i] = cstat_data[,i]/sum(cstat_data[,i])
    }  

# Create Distance Matrix
#samplePA.dist <- vegdist(decostand(t(dataPA),method="log"),method="bray")
sampleREL.dist <- vegdist(decostand(t(dataREL),method="log"),method="bray")
#sampleREL.dist <- vegdist(t(dataREL),method="bray")

# Principal Coordinates Analysis
cstat_pcoa <- cmdscale(sampleREL.dist,k=3,eig=TRUE,add=FALSE) 
  # Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
    	# eig=TRUE returns eigenvalues; k = # of dimensions to calculate

# Percent Variance Explained Using PCoA (Axis 1,2,3)
explainvar1 <- round(cstat_pcoa$eig[1]/sum(cstat_pcoa$eig)*100,1) 
explainvar2 <- round(cstat_pcoa$eig[2]/sum(cstat_pcoa$eig)*100,1)
explainvar3 <- round(cstat_pcoa$eig[3]/sum(cstat_pcoa$eig)*100,1)
  
pcoap <- merge(as.data.frame(cstat_pcoa$points),design,by=0,all.x=T)
rownames(pcoap) <- rownames(cstat_pcoa$points)
#write.csv(file="./supporting_files/data/pcoap",pcoap)
#pcoap2 <- subset(pcoap,pcoap$limitation == "T")

pdf(file = "./supporting_files/figure_betaPCOA-all.pdf", width = 5, height = 5, pointsize = 10)

# Plot Parameters
par(mar=c(5,5,1,1))#, oma=c(3,1,1,1)+0.1 )
x.dim <- c(min(pcoap$V1)-(max(pcoap$V1)*0.1),max(pcoap$V1)+(max(pcoap$V1)*0.1))
y.dim <- c(min(pcoap$V2)-(max(pcoap$V2)*0.1),max(pcoap$V2)+(max(pcoap$V2)*0.1))

# Initiate Plot
plot(pcoap$V1, pcoap$V2, xlab=paste("PCoA Axis 1 (",explainvar1, "%)", sep="")
    , ylab=paste("PCoA Axis 2 (",explainvar2, "%)", sep=""), 
    xlim=x.dim, ylim=y.dim, pch=16, cex=2.0, type="n",xaxt="n",
    yaxt="n", cex.lab=1.5, cex.axis=1.2)
axis(side=1, las=1)   
axis(side=2, las=1)    
abline(h=0, lty="dotted")  
abline(v=0, lty="dotted")
box(lwd=2)

# Make Plot Symbols in PCoA Reflect Treatment
 
mol.shape <- rep(NA, dim(pcoap)[1])
  for (i in 1:length(mol.shape)){
    if (pcoap$treatment[i] == "T"){mol.shape[i] = 21}
    else {mol.shape[i] = 22}
	}

slope.color <- rep(NA, dim(pcoap)[1])
  for (i in 1:length(slope.color)){
    if (pcoap$limitation[i] == "N") {slope.color[i] = "black"}
    else {slope.color[i] = "white"}  
	} 

# Add Points & Ellipses
points(pcoap$V1, pcoap$V2, pch = mol.shape, cex = 2.5, 
       bg = slope.color, col = "gray25")
#points(pcoap$V1, pcoap$V2, pch = mol.shape, cex = (pcoap$time+1)*0.2, 
#       bg = slope.color, col = "gray25")
#text(pcoap$V1, pcoap$V2, labels = pcoap$time, pos = 3)  
#text(pcoap$V1, pcoap$V2, labels = pcoap$cID, pos = 1) 

dev.off()

#p = ggplot(pcoap,aes(x=V1, y = V2, shape = treatment )) + geom_point(size =4) + theme(legend.position="none")
#p = p + facet_grid(~cID) +  
#  geom_path(aes(x = V1, y = V2, size = time/1.5), color = "grey80") +
#  geom_point(aes(size = time), color = "black") + xlab("Axis.1") + ylab("Axis.2") 

#print(p)

#ggsave(filename = "./figure_cIDbeta.pdf", plot = p, height = 2.5, width = 10, units = "in", scale = 2)

```

```{r 2.2_PERMANOVA-all}
#BC = t(dataREL)
#BC = vegdist(t(dataREL),method="bray")
#df =  design
#results3 <- adonis(BC ~ time*treatment*limitation, data = df, 
#                  strata = df$cID, permutations = 999, "bray")

cstat.adonis.data <- t(dataREL)
cstat.adonis.data <- merge(design,cstat.adonis.data, by=0,all.x=T)
cstat.adonis.data <- cstat.adonis.data[-17,]
#cstat.adonis.data <- subset(cstat.adonis.data,cstat.adonis.data$trt == "T")
#cstat.adonis.data <- cstat.adonis.data[,1:36]

results3 <- adonis(cstat.adonis.data[,6:ncol(cstat.adonis.data)] ~ 
         limitation*treatment*time, strata = cstat.adonis.data$cID, data = cstat.adonis.data,
       method="bray", permutations=999)


```

####Observations
Time, nutrient limitation and exposure to phage signficantly affected the microbial community within each chemostat (PERMANOVA, time + limitation*phage, p < 0.001)

```{r PCoA-lim}
pcoap <- subset(pcoap,pcoap$treatment == "T")

pdf(file = "./supporting_files/figure_betaPCOA-lim.pdf", width = 5, height = 5, pointsize = 10)

# Plot Parameters
par(mar=c(5,5,1,1))#, oma=c(3,1,1,1)+0.1 )
x.dim <- c(min(pcoap$V1)-(max(pcoap$V1)*0.1),max(pcoap$V1)+(max(pcoap$V1)*0.1))
y.dim <- c(min(pcoap$V2)-(max(pcoap$V2)*0.1),max(pcoap$V2)+(max(pcoap$V2)*0.1))

# Initiate Plot
plot(pcoap$V1, pcoap$V2, xlab=paste("PCoA Axis 1 (",explainvar1, "%)", sep="")
    , ylab=paste("PCoA Axis 2 (",explainvar2, "%)", sep=""), 
    xlim=x.dim, ylim=y.dim, pch=16, cex=2.0, type="n",xaxt="n",
    yaxt="n", cex.lab=1.5, cex.axis=1.2)
axis(side=1, las=1)   
axis(side=2, las=1)    
abline(h=0, lty="dotted")  
abline(v=0, lty="dotted")
box(lwd=2)

# Make Plot Symbols in PCoA Reflect Treatment
 

mol.shape <- rep(NA, dim(pcoap)[1])
  for (i in 1:length(mol.shape)){
    if (pcoap$limitation[i] == "N"){mol.shape[i] = 21}
    else {mol.shape[i] = 22}
	}

# Add Points & Ellipses
points(pcoap$V1, pcoap$V2, pch = mol.shape, cex = 2.5, 
       bg = c("white","grey25","grey50","grey75","black"), col = "gray25")
#points(pcoap$V1, pcoap$V2, pch = mol.shape, cex = (pcoap$time+1)*0.2, 
#       bg = slope.color, col = "gray25")
#text(pcoap$V1, pcoap$V2, labels = pcoap$time, pos = 3)  
#text(pcoap$V1, pcoap$V2, labels = pcoap$cID, pos = 1) 

dev.off()
```

```{r 2.2_PERMANOVA-lim+Ph}
#BC = t(dataREL)
#BC = vegdist(t(dataREL),method="bray")
#df =  design
#results3 <- adonis(BC ~ time*treatment*limitation, data = df, 
#                  strata = df$cID, permutations = 999, "bray")

cstat.adonis.data <- t(dataREL)
designT <- design[design$treatment == "T",]
cstat.adonis.data <- merge(designT,cstat.adonis.data, by=0,all.x=T)
cstat.adonis.data <- cstat.adonis.data[-12,]


results3 <- adonis(cstat.adonis.data[,6:ncol(cstat.adonis.data)] ~ 
         limitation*day, strata = cstat.adonis.data$cID, data = cstat.adonis.data,
       method="bray", permutations=999)


```

```{r 2.2_PERMANOVA-lim-Ph}
#BC = t(dataREL)
#BC = vegdist(t(dataREL),method="bray")
#df =  design
#results3 <- adonis(BC ~ time*treatment*limitation, data = df, 
#                  strata = df$cID, permutations = 999, "bray")

cstat.adonis.data <- t(dataREL)
designC <- design[design$treatment == "C",]
cstat.adonis.data <- merge(designC,cstat.adonis.data, by=0,all.x=T)


results3 <- adonis(cstat.adonis.data[,6:ncol(cstat.adonis.data)] ~ 
         limitation*day, strata = cstat.adonis.data$cID, data = cstat.adonis.data,
       method="bray", permutations=999)


```


```{r 2.2_PERMANOVA-NL}
#BC = t(dataREL)
#BC = vegdist(t(dataREL),method="bray")
#df =  design
#results3 <- adonis(BC ~ time*treatment*limitation, data = df, 
#                  strata = df$cID, permutations = 999, "bray")

cstat.adonis.data <- t(dataREL)
designNL <- design[design$limitation == "N",]
cstat.adonis.data <- merge(designNL,cstat.adonis.data, by=0,all.x=T)


results3 <- adonis(cstat.adonis.data[,6:ncol(cstat.adonis.data)] ~ 
         treatment*day, strata = cstat.adonis.data$cID, data = cstat.adonis.data,
       method="bray", permutations=999)


```

```{r 2.2_PERMANOVA-PL}
#BC = t(dataREL)
#BC = vegdist(t(dataREL),method="bray")
#df =  design
#results3 <- adonis(BC ~ time*treatment*limitation, data = df, 
#                  strata = df$cID, permutations = 999, "bray")

cstat.adonis.data <- t(dataREL)
designPL <- design[design$limitation == "P",]
cstat.adonis.data <- merge(designPL,cstat.adonis.data, by=0,all.x=T)
cstat.adonis.data <- cstat.adonis.data[-9,]

results3 <- adonis(cstat.adonis.data[,6:ncol(cstat.adonis.data)] ~ 
         treatment*day, strata = cstat.adonis.data$cID, data = cstat.adonis.data,
       method="bray", permutations=999)


```

```{r 2.2_temporal_composition, echo = FALSE, fig.width=7.5, fig.height=4}
library(ggplot2)
library(ggthemes)
library(reshape2)

# Select most abundant taxa with synechococcus (OTU3); adjust number of reads by mean copy number for each strain
cpNo <- c(3.57,2.33,1.85,3.08,5.33,4.86,2.50,1)
cstat_data_top <- as.data.frame(t(cstat_data[1:7,]))
cstat_data_top$others <- NA

for(i in 1:nrow(cstat_data_top)){
  cstat_data_top[i,8] <- rowSums(t(abs(cstat_data[8:nrow(cstat_data),])))[[i]]
}  


for(i in 1:length(cpNo)){
  cstat_data_top[,i] <- cstat_data_top[,i]/cpNo[i]
}

cstat_data_top <- t(cstat_data_top)

# With Syn
dataRELtop <- cstat_data_top
  for(i in 1:ncol(cstat_data_top)){
    dataRELtop[,i] = cstat_data_top[,i]/sum(cstat_data_top[,i])
    } 

top.abd.design <- merge(design,t(dataRELtop), by=0,all.x=T)

colnames(top.abd.design) <- c(colnames(top.abd.design)[1:6],"Sulfitobacter","Sphingomonas", "Synechococcus",
                              "Rhizobium","Arthrobacter","Alcanivorax","Pseudomonas","others")
melt.abd <- melt(top.abd.design, id=c("time","day","cID","Row.names","limitation","treatment"))
melt.abd$day <- as.numeric(as.character(melt.abd$day))



tes <- melt.abd[-which(melt.abd$cID == "PL3"),]
colnames(tes)<- c(colnames(tes)[1:6],"Taxa","value")
p <- ggplot(tes,aes(x = day, y = value, fill = Taxa)) + 
  geom_area(position = 'stack', colour ="black") + 
  facet_wrap(~cID, ncol = 4) +
  #facet_gridtreatment ~ limitation) +
  labs(x = "Time (days)", y = "Relative Abundance")+
  #scale_fill_grey(start = 0 , end = 0.9)
  scale_fill_brewer(palette = "Greens")
  
#p$data$Taxa <- factor(p$data$Taxa, levels = #c("Pseudomonas","Alcanivorax","Arthrobacter","Rhizobium","Sphingomonas","Sulfitobacter"))

print(p)

#ggsave(filename="./figure_taxarelabd.pdf", plot = p,width = 8, height = 4, units = "in",scale = 1.5)

# Without Syn
cstat_data_top_wosyn <- cstat_data_top[-3,]
dataRELtop_wo <- cstat_data_top_wosyn
  for(i in 1:ncol(cstat_data_top_wosyn)){
    dataRELtop_wo[,i] = cstat_data_top_wosyn[,i]/sum(cstat_data_top_wosyn[,i])
    } 
relAbd <- merge(design,t(dataRELtop_wo), by=0,all.x=T)
relAbdflow <- merge(relAbd,flow.counts, by.x = "Row.names", by.y = "seq.ID")
relAbdflow <- relAbdflow[,c(1:13,18:19)]

for(i in 1:nrow(relAbdflow)){
  relAbdflow[i,7:13] <- relAbdflow[i,7:13]*relAbdflow[i,15]
}

ggplot(data = relAbdflow, aes(x = day, y = flow.syn)) +
  facet_wrap(~cID.x) +
  geom_path()+
  geom_line(aes(y=Otu001), color = "red", size = 2) +
  geom_line(aes(y=Otu002), color = "blue", size = 2) +
  geom_line(aes(y=Otu004), color = "green", size = 2)+
  geom_line(aes(y=Otu005), color = "orange", size = 2)+
  geom_line(aes(y=Otu006), color = "purple", size = 2)+
  geom_line(aes(y=Otu007), color = "yellow", size = 2)

```

```{r perm-tax}

dat = p$data
dat$time = as.factor(dat$time)
for(i in unique(dat$Taxa)){
  tmpdat = dat[dat$Taxa == i,]
  
  fit <- lme(value ~ limitation * time * treatment, random = ~ 1 | cID,
                 correlation = corAR1(form = ~ 1 | cID),
                 data = tmpdat, na.action = na.exclude)
  fit.dat = anova.lme(fit, type = "marginal", adjustSigma = F)
  print(i)
  print(fit.dat)
  
}

# calculate the mean relative densities for each of the major taxa for each treatment
dat.rhi <- dat[dat$Taxa == "Rhizobium",]
aggregate(value~ limitation + treatment, data = dat.rhi, mean)
aggregate(value~ cID, data = dat.rhi, mean, na.action = na.exclude)
aggregate(value~ limitation, data = dat.rhi, mean)
aggregate(value~ limitation, data = dat.rhi, sd)

```

```{r}

#install.packages("Hmisc")
require(Hmisc)

mat.N <- as.matrix(top.abd.design[,7:13][top.abd.design$limitation=="N",])
mat.P <- as.matrix(top.abd.design[,7:13][top.abd.design$limitation=="P",])
mat.P <- na.omit(mat.P)
ts.cor.N <- rcorr(mat.N,type="pearson")
ts.cor.P <- rcorr(mat.P,type="pearson")

syn.N <- mat.N[,3]
syn.P <- mat.P[,3]
sulf.N <- mat.N[,1]
sph.N <- mat.N[,2]
ath.N <- mat.N[,5]
ath.N <- mat.N[,5]

for(i in unique(top.abd.design$cID)[-9]){
  tmp <- as.matrix(top.abd.design[top.abd.design$cID == i,7:13])
  tmp.cor <- rcorr(tmp,type="pearson")
  print(i)
  print(tmp.cor)
}
```
```{r}
pdf(file = "./figure_NLccf.pdf", width = 5, height = 4.5, pointsize= 10)

par(mfrow = c(2,4))
for(i in 1:7){
  ccf(syn.N,mat.N[,i], main = colnames(mat.N)[i])
}

dev.off()
#Sphingomonas = out of phase, Rhizobium out of phase, Athrobacter
```

```{r}
pdf(file = "./figure_PLccf.pdf", width = 5, height = 4.5, pointsize = 10)


par(mfrow = c(2,4))
for(i in 1:7){
  ccf(syn.P,mat.P[,i], main = colnames(mat.P)[i])
}

dev.off()
#Sulfitobacter out of phase. lag -2, Rhizobium, 
```

Abundances of *Sphingomonas* (*r* = -0.73, p-value = 0.0003) and *Rhizobium* (*r* = -0.56, p-value = 0.0103) were significantly correlated with *Synechococcus* under N-limitation. *Sphingomonas* was also significantly correlated
Abundances of *Sulfitobacter* (*r* = -0.44, p-value = 0.036) and *Alcanivorax* (*r* = -0.42, p-value = 0.049) were significantly correlated with *Synechococcus* densities under P-limitation.

Need to make table with syn mean, phage mean, het mean and the rel mean abundances of each taxa within the community

# Phyloseq analysis
```{r 2.1_section_data}
library(phyloseq)

# Set ggplot theme
theme_set(theme_bw())

# OTU and Taxonomy Files
mothtax <- "./mothur/20151029/Cstat_all.final.an.unique.cons.taxonomy"
mothshare <- "./mothur/20151029/Cstat_all.final.an.shared"
#mothtree <- "./mothur/20151029/Cstat_all.final.an.tree"
cutoff <- "unique"

# Create phyloseq object

physeq <- import_mothur(mothur_shared_file = mothshare, 
                        mothur_constaxonomy_file = mothtax,
                        #mothur_tree_file = mothtree,
                        cutoff)

# Add sample information
# Sample Data
design.in <- "./data/design.txt"
design <- read.delim(design.in, header=T, row.names=1)
samp.data <- sample_data(design)
physeq <- merge_phyloseq(physeq, samp.data)

physeq
otu_table(physeq)[1:5,1:5]
tax_table(physeq)[1:5,1:5]
sample_data(physeq)[1:10]
taxa_names(physeq)[1:10]

```

```{r 2.1_explore_plots, include=FALSE}
# Plot tree for top 10 taxa in data set
myTaxa = names(sort(taxa_sums(physeq), decreasing = TRUE)[1:10])
ex1 = prune_taxa(myTaxa, physeq)

# Phylogenetic tree by chemostat
#plot_tree(ex1, color = "cID", label.tips = "Phylum", 
#          ladderize = "left", justify = "left" , size = "Abundance")
```

```{r 2.1_preprocessing}
# Remove taxa not seen more than 2 times in at least 20% of the samples. This should protect against an OTU with a small mean.
physeq.trim = filter_taxa(physeq, function(x) sum(x > 2) > (0.2*length(x)), TRUE)

#Is there a need to standardize the data by median sequencing depth?

```

##Alpha Diversity
```{r 2.1_alpha_diversity, echo = FALSE, include = FALSE}
# Alpha Diversity 
p = plot_richness(physeq.trim, x = "day", 
                  measures = "Shannon", col = "cID", shape = "treatment")
p1 = p + geom_point(cex = 5)+ geom_path()+ facet_grid(treatment~limitation) + theme(axis.text.x=element_text(angle = 0, hjust = 0.5))
print(p1)

ggsave(filename = "./figure_alpha-time.pdf", plot = p1, width = 5, height = 4.5, units = "in", dpi = 2400, scale = 1.75)

p2 = plot_richness(physeq.trim, x = "treatment", 
                  measures = c("Observed","Shannon"), 
                  col = "limitation")+ geom_boxplot()+geom_point()
#print(p2)
```

```{r RMANOVA-alphaDiv}
ar1.alpha <- lme(value ~ limitation * time * treatment, random = ~ 1 | cID, 
            correlation = corAR1(form = ~ 1 | cID),
            data = p1$data)
#arma.het <- lme(flow.het ~ lim * time * trt, random = ~ 1 | cID, 
#            correlation = corARMA(form = ~ 1 , p = 1,q = 2),
#            data = flow.counts)

#anova(ar1.het,arma.het)
summary(ar1.het)
anova.lme(ar1.alpha, type = "marginal", adjustSigma = F)
```

###2.1.2 Beta Diversity 
```{r 2.1_beta_diversity, echo = FALSE, include = FALSE, eval = FALSE}
# Beta Diversity
# normalize data for the most abundant taxa?
# physeq.trim.tr = transform_sample_counts(physeq.trim, function(x) decostand(x))

cstat_pcoa = ordinate(physeq.trim, method = "PCoA", distance = "bray")
shapes = c(0,1,2,3,4,5,6,7,8,9)
p = plot_ordination(physeq.trim, cstat_pcoa, shape = "treatment", color = "limitation") 
p + facet_grid(~cID) +  
  geom_path(aes(x = Axis.1, y = Axis.2, size = time/1.5), color = "grey80") +
  geom_point(aes(size = time), color = "black") 

print(p)

```

```{r 2.1_PERMANOVA, warning=FALSE, echo = FALSE}
require(vegan)
df = data.frame(sample_data(physeq.trim))
BC = distance(physeq.trim, "bray")

results = adonis(BC ~ limitation*treatment*time, strata = df$cID,
                 data = df, permutations = 999)


OTU = t(as(otu_table(physeq.trim), "matrix"))
results2 = adonis(OTU ~ time*limitation*treatment , strata= df$cID,
                  data = df, permutations = 999)

```
